---
title: "Plotting figures for Beetle flower visitor review"
output:
  html_notebook:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
---

In this notebook, we will plot the results of our literature summaries made with Anthropic Sonnet 3.5 API, and also the beetle phylogeny highlighting beetle flower visitors and pollinators.

Read packages.
```{r}
library(tidyverse)
library(jsonlite)
library(sf)
library(rnaturalearth)
library(ggthemes)
library(patchwork)
library(cowplot)
library(reticulate)
library(httr)
library(jsonlite)
library(treeio)
library(deeptime)
library(ape)
library(ggtree)
library(phytools)
library(vegan)
library(ggrepel)

use_condaenv("ARE_review")
rm(list=ls())
```

# Preprocessing and filtering data

In this first part of the notebook, we will

1. Read the json file with results from Sonnet 3.5 anaylses
2. Reconcile geographical locations to those in [Natural Earth](https://www.naturalearthdata.com/) using the `rnaturalearth` package
3. Reconcile plant species and higher taxon names to the classification of [World Flora Online](https://www.worldfloraonline.org/)


## Parse JSON output
Let's read the json output:
```{r}
review_data = fromJSON("analysis/flower_visitor_records.json")
```

Transforming into a data frame with some error handling:
```{r}
# Create a safe version of our processing function
safe_process <- possibly(
  function(bibtex_id) {
    element <- review_data[[bibtex_id]]
    
    if (!is.null(element$json) && 
        !is.null(element$json$has_primary_visitor_data) &&
        isTRUE(element$json$has_primary_visitor_data) &&  # using isTRUE is safer
        !is.null(element$json$records)) {
      
      element$json$records %>%
        mutate(bibtex_id = bibtex_id)
    }
  },
  otherwise = NULL  # Return NULL if anything goes wrong
)

# Apply the safe function
combined_records <- names(review_data) %>%
  map_dfr(safe_process)

combined_records
```
## Reconcile geograpghy

Country names are sometimes not standardized. Let's have a look at the distribution of country names in our dataset:
```{r}
combined_records$country %>% table %>% sort(decreasing = TRUE)
```

Let's now add ISO codes to standardize country names. Records that cannot be mapped to a single existing country will receive NA and be ignored from here on:

```{r}
country_codes <- tribble(
  ~country, ~country_ISO_code,
  "Costa Rica", "CRI",
  "France", "FRA",
  "Spain", "ESP",
  "Australia", "AUS",
  "USA", "USA",
  "United States", "USA",
  "United States of America", "USA",
  "Peru", "PER",
  "Ecuador", "ECU",
  "Colombia", "COL",
  "Ecuador and Peru", NA,
  "Mexico", "MEX",
  "México", "MEX",
  "Brazil", "BRA",
  "Thailand", "THA",
  "Venezuela", "VEN",
  "French Guiana", "FR-GF",
  "Argentina", "ARG",
  "Gabon", "GAB",
  "Canada", "CAN",
  "South Africa", "ZAF",
  "Republic of South Africa", "ZAF",
  "Sweden", "SWE",
  "Malaysia", "MYS",
  "Japan", "JPN",
  "Pakistan", "PAK",
  "Germany", "DEU",
  "Puerto Rico", "PRI",
  "Chile", "CHL",
  "Madagascar", "MDG",
  "China", "CHN",
  "Denmark", "DNK",
  "Benin", "BEN",
  "Iran", "IRN",
  "Morocco", "MAR",
  "Jordan", "JOR",
  "Czech Republic", "CZE",
  "Belgium", "BEL",
  "United Kingdom", "GBR",
  "UK", "GBR",
  "Great Britain", "GBR",
  "England", "GBR",
  "Scotland", "GBR",
  "Ireland", "IRL",
  "Saudi Arabia", "SAU",
  "Panama", "PAN",
  "Italy", "ITA",
  "Poland", "POL",
  "Finland", "FIN",
  "Suriname", "SUR",
  "India", "IND",
  "Tanzania", "TZA",
  "Trinidad and Tobago", "TTO",
  "Trinidad", "TTO",
  "Jamaica", "JAM",
  "Israel", "ISR",
  "Bolivia", "BOL",
  "Turkey", "TUR",
  "Türkiye", "TUR",
  "Cameroon", "CMR",
  "Ghana", "GHA",
  "American Samoa", "ASM",
  "Cuba", "CUB",
  "Multiple European countries and USA", NA,
  "Algeria", "DZA",
  "Zimbabwe", "ZWE",
  "El Salvador", "SLV",
  "New Zealand", "NZL",
  "Netherlands", "NLD",
  "The Netherlands", "NLD",
  "Mauritius", "MUS",
  "Democratic Republic of the Congo", "COD",
  "Democratic Republic of Congo", "COD",
  "Namibia", "NAM",
  "Greece", "GRC",
  "none", NA,
  "Botswana", "BWA",
  "Uruguay", "URY",
  "USA and Canada", NA,
  "Korea", "KOR",
  "South Korea", "KOR",
  "Lithuania", "LTU",
  "Taiwan", "TWN",
  "Cambodia", "KHM",
  "Hong Kong", NA,
  "Portugal", "PRT",
  "Indonesia", "IDN",
  "Austria", "AUT",
  "Comoros", "COM",
  "Zambia", "ZMB",
  "Samoa", "WSM",
  "Western Samoa", "WSM",
  "Switzerland", "CHE",
  "Burkina Faso", "BFA",
  "Cyprus", "CYP",
  "Nepal", "NPL",
  "Senegal", "SEN",
  "Norway", "NOR",
  "Papua New Guinea", "PNG",
  "Nigeria", "NGA",
  "Multiple European countries", NA,
  "Russia", "RUS",
  "New Caledonia", "NCL",
  "Seychelles", "SYC",
  "Cabo Verde", "CPV",
  "Cape Verde", "CPV",
  "Egypt", "EGY",
  "Ukraine", "UKR",
  "Singapore", "SGP",
  "Guinea", "GIN",
  "Nicaragua", "NIC",
  "Brunei", "BRN",
  "Brunei Darussalam", "BRN",
  "Fiji", "FJI",
  "Ethiopia", "ETH",
  "French Polynesia", "PYF",
  "Ivory Coast", "CIV",
  "Spain and Portugal", NA,
  "Vietnam", "VNM",
  "Vanuatu", "VUT",
  "Malawi", "MWI",
  "Lesotho", "LSO",
  "Kenya", "KEN",
  "Laos", "LAO",
  "Canada and USA", NA,
  "Bulgaria", "BGR",
  "The Gambia", "GMB",
  "Dominica", "DMA",
  "Guatemala", "GTM",
  "Montenegro", "MNE",
  "Serbia", "SRB",
  "Dominican Republic", "DOM",
  "Reunion Island", "FR-RE",
  "Reunion", "FR-RE",
  "Réunion", "FR-RE",
  "Hungary", "HUN",
  "Bangladesh", "BGD",
  "Lebanon", "LBN",
  "Philippines", "PHL",
  "Estonia", "EST",
  "Croatia", "HRV",
  "Andorra", "AND",
  "Paraguay", "PRY",
  "Uganda", "UGA",
  "Saint Helena", "SHN",
  "Mayotte", "FR-YT",
  "Bahamas", "BHS",
  "Sri Lanka", "LKA",
  "Guyana", "GUY",
  "Georgia", "GEO",
  "Crimea", NA,
  "Bosnia and Herzegovina", "BIH",
  "United States and Mexico", NA,
  "French Southern and Antarctic Lands", "ATF",
  "Uzbekistan", "UZB"
)

combined_records = combined_records %>% left_join(country_codes)

combined_records
```
Now let's combine this with data. For countries with small latitudinal range, we will combine data by country. For countries with large latitudinal range, we will do it by states. Let's get country-level data first:
```{r}
# Get world map data from rnatural eath
world_sf <- ne_countries(scale = "medium", returnclass = "sf")
```

let's now sort countries by latitudinal range.
```{r}
country_lat_range = world_sf %>% mutate(
    min_lat = map_dbl(geometry, ~ st_bbox(.x)[["ymin"]]),
    max_lat = map_dbl(geometry, ~ st_bbox(.x)[["ymax"]]),
    lat_range = max_lat - min_lat
  ) %>%
  select(name, adm0_a3, min_lat, max_lat, lat_range) %>%
  arrange(desc(lat_range)) %>%
  as_tibble()

country_lat_range
```

Let's split records by state for all countries with more than 20 degrees of latitudinal range. Let's find the states for these countries. First, let's figure out which states from our literature database don't map to states in natural earth data:

```{r}
higher_lat_range_codes = country_lat_range %>% filter(lat_range > 20) %>% pull(adm0_a3)

existing_states = ne_states() %>%
  filter(adm0_a3 %in% higher_lat_range_codes) %>%
  as_tibble() %>%
  select(state_name = name,country_name = admin, country_ISO_code=adm0_a3, state_ISO_code = iso_3166_2)
  
combined_records %>% filter(country_ISO_code %in% higher_lat_range_codes) %>%
  select(country,state_name = state_province,country_ISO_code) %>%
  distinct() %>%
  left_join(existing_states,by = c('state_name','country_ISO_code')) %>%
  filter(is.na(country_name)) %>%
  select(-country_name)
```

We will now use Claude to map these states to names that exist in the natural earth database. Let's save a table to use as project knowledge.

```{r}
existing_states %>% write_csv("~/Downloads/ne_states.csv")
```

This is the mapping that Claude generated, with some manual corrections:
```{r}
# Create mapping table for state names to ISO codes
state_iso_mapping <- data.frame(
  original_state_name = c(
    "Occitanie", "Southern Cape", "Ibaraki Prefecture", "Tokyo Prefecture", "V Region",
    "Cordillera Province", "Saint-Philippe", "Brittany", "Valparaíso Region", "Juan Fernández Islands",
    "Santiago Metropolitan Region", "Metropolitan Region", "North West Province", "Languedoc-Roussillon", "Quebec",
    "Shandong Province", "Yunnan Province", "Eastern Cape Province", "Region of Coquimbo", "Valdivia",
    "Multiple states", "Idaho, Nevada, Utah", "California, Idaho, Nevada, Oregon, Utah", "Hokkaido", "Goias",
    "Midi-Pyrénées", "Inner Mongolia", "Rio Negro", "Gansu Province", "Guangxi Province",
    "Jiangxi Province", "Eastern Cape and KwaZulu-Natal", "Los Rios", "Los Ríos Region", "Goiás State",
    "North Island", "Northern Cape Province", "Minais Gerais", "Pyrenees", "Rio Negro and Neuquen",
    "Federal District", "Shizuoka Prefecture", "Kyoto Prefecture", "Kyoto", "Maule Region",
    "Northern Cape/Western Cape", "Hyogo", "Altai Republic", "Krasnodar Territory", "Cordoba",
    "Salta/Jujuy", "Fukuoka Prefecture", "Kochi Prefecture", "Hyogo Prefecture", "São Paulo/Rio de Janeiro border",
    "Henan Province", "Liaoning Province", "Entre Rios", "Central Chile", "Kerala and Tamil Nadu",
    "Western Australia and Northern Territory", "Tibet", "Hong Kong", "Jiangsu Province", "Honshu",
    "Nagano Prefecture", "KwaZulu Natal", "Para", "Yamanashi Prefecture", "Maharashtra and Goa",
    "Piaui", "Fujian Province", "Hubei Province", "Shannxi Province", "Valparaiso",
    "Qinghai Province", "Rhône-Alpes", "Mie Prefecture", "Kagoshima Prefecture", "Nara Prefecture",
    "Miyagi Prefecture", "Miyagi Prefecture and Yamagata Prefecture", "Region Metropolitana", "Biobío", "Aomori Prefecture",
    "North-West Province", "Natal", "Sichuan Province", "Western Cape Province", "Cape Province", 
    "Wakayama Prefecture", "Taiwan", "Ishikawa Prefecture", "Okinawa Prefecture", "Amami Islands",
    "Ryukyu Islands", "Yakushima", "Toyama Prefecture", "South Holland", "Kyushu",
    "Tarapacá Region", "Crimean Peninsula", "Limpopo Province", "Okayama Prefecture", "Kerala and Karnataka",
    "Burgundy", "Nouvelle-Aquitaine", "Réunion", "South Holland", "Heilongjiang Province",
    "Guangdong Province", "Moscow Oblast", "Tochigi Prefecture", "Xinjiang Uyghur Autonomous Region", "Guam",
    "Shiga Prefecture", "IV Región", "Gifu Prefecture", "Hyogo, Okayama, and Tottori Prefectures", "Multiple prefectures",
    "Tucuman", "Kagawa Prefecture", "Colchagua", "Talca", "South Island",
    "Ryukyu Archipelago", "Miyazaki Prefecture", "Qinghai-Tibet", "Shimane Prefecture", "Primorsky Krai", 
    "Western Cape and Eastern Cape", "Bonin Islands", "Hokkaido Prefecture", "Kinki", "Osaka Prefecture",
    "Akita Prefecture", "Rio de Janeiro/Minas Gerais", "Orissa", "Auckland Region", "Île-de-France",
    "Washington and Oregon", "Multiple including Gauteng and KwaZulu-Natal", "Banks Peninsula", "Guangxi Zhuang Autonomous Region", "Corsica",
    "Region of Maule", "Victoria and New South Wales", "Nelson", "Chiloé", "VIII Region",
    "Washington and Idaho", "Idaho and Washington", "Oita Prefecture", "Oregon/Washington/Idaho", "Western Ghats",
    "Hiroshima Prefecture", "Sikkim and Darjeeling", "Maule and Los Ríos Regions", "Altai Krai", "Provence",
    "Provence-Alpes-Côte d'Azur", "Chiba Prefecture", "Osaka", "Kumamoto Prefecture", "Nagasaki Prefecture",
    "Ogasawara Islands", "Tottori Prefecture", "Yamaguchi Prefecture", "Ryukyu archipelago", "Saga Prefecture",
    "Aichi Prefecture", "Shikoku", "Shanxi Province", "Hunan Province", "Región Metropolitana",
    "Santiago", "Shaanxi Province", "Hyogo, Osaka, Nara", "Moscow oblast", "Tula oblast",
    "Región de Los Lagos", "Espirito Santo", "Region V", "Region La Araucania", "Valparaiso Region",
    "multiple states", "Jammu", "Region of Bernardo O'Higgins", "Inner Mongolia Autonomous Region", "Languedoc",
    "Zhejiang Province", "Xisha Islands", "Alabama, Georgia, South Carolina, North Carolina", "Honshu and Izu Islands", "Tokushima Prefecture",
    "Xinjiang Uygur Autonomous Region", "California/Oregon border", "Mpumulanga"
  ),
  state_ISO_code = c(
    "FRA", NA, "JP-08", "JP-13", NA,  # 1-5
    NA, "FR-RE", "FRA", NA, NA,      # 6-10
    "CL-RM", "CL-RM", "ZA-NW", NA, "CA-QC",  # 11-15
    "CN-SD", "CN-YN", "ZA-EC", "CL-CO", NA,  # 16-20
    NA, NA, NA, "JP-01", "BR-GO",        # 21-25
    NA, "CN-NM", "AR-R", "CN-GS", "CN-GX",  # 26-30
    "CN-JX", NA, "CL-LR", "CL-LR", "BR-GO",  # 31-35
    NA, "ZA-NC", "BR-MG", NA, NA,        # 36-40
    "BR-DF", "JP-22", "JP-26", "JP-26", "CL-ML",  # 41-45
    NA, "JP-28", "RU-AL", "RU-KDA", "AR-X",  # 46-50
    NA, "JP-40", "JP-39", "JP-28", NA,   # 51-55
    "CN-HA", "CN-LN", "AR-E", NA, NA,    # 56-60
    NA, "CN-XZ", "CN-GD", "CN-SC", NA,   # 61-65
    "JP-20", "ZA-NL", "BR-PA", "JP-19", NA,  # 66-70
    "BR-PI", "CN-FJ", "CN-HB", NA, "CL-VS",  # 71-75
    "CN-QH", NA, "JP-24", "JP-46", "JP-29",  # 76-80
    "JP-04", NA, "CL-RM", "CL-BI", "JP-02",  # 81-85
    "ZA-NW", NA, "CN-SC", "ZA-WC", NA,   # 86-90
    "JP-30", "TWN", "JP-17", "JP-47", NA,  # 91-95
    NA, NA, "JP-16", "NL-ZH", NA,        # 96-100
    "CL-TA", NA, "ZA-LP", "JP-33", NA,   # 101-105
    "FRA", "FRA", "FR-RE", "NL-ZH", "CN-HL",  # 106-110
    "CN-GD", "RU-MOS", "JP-09", "CN-XJ", "GU-X01~",  # 111-115
    "JP-25", NA, "JP-21", NA, NA,        # 116-120
    "AR-T", "JP-37", NA, NA, NA,         # 121-125
    NA, "JP-45", NA, "JP-32", "RU-PRI",  # 126-130
    NA, NA, "JP-01", NA, "JP-27",        # 131-135
    "JP-05", NA, "IN-OR", "NZ-X06~", "FRA",  # 136-140
    NA, NA, NA, "CN-GX", "FRA",       # 141-145
    "CL-ML", NA, "NZ-NSN", NA, NA,       # 146-150
    NA, NA, "JP-44", NA, NA,             # 151-155
    "JP-34", NA, NA, "RU-ALT", NA,       # 156-160
    "FRA", "JP-12", "JP-27", "JP-43", "JP-42",  # 161-165
    NA, "JP-31", "JP-35", NA, "JP-41",   # 166-170
    "JP-23", NA, "CN-SX", "CN-HN", "CL-RM",  # 171-175
    "CL-RM", "CN-SN", NA, "RU-MOS", "RU-TUL",  # 176-180
    "CL-LL", "BR-ES", NA, "CL-AR", "CL-VS",  # 181-185
    NA, "IN-JK", "CL-LI", "CN-NM", NA,   # 186-190
    "CN-ZJ", NA, NA, NA, "JP-36",        # 191-195
    "CN-XJ", NA, "ZA-MP"                 # 196-198
  )
)

state_iso_mapping = state_iso_mapping %>% filter(!is.na(state_ISO_code))
```


After this first step, I asked claude to incorporate locality information and see if we can map a few more:
```{r}
additional_state_mapping = data.frame(
  original_state_name = c(
    "Southern Cape", # Localities: Vasselot Nature Reserve, Kranshoek, Harkerville, Midway
    "V Region",      # Multiple localities all in Valparaíso Region
    "Cordillera Province", # El Volcán, Farellones, Termas Colina
    "Valparaíso Region", # Includes Juan Fernández Islands localities
    "Juan Fernández Islands", # Part of Valparaíso Region
    "Languedoc-Roussillon", # Multiple localities but historical region
    "Valdivia",    # Specific locality: Botanical Garden of Universidad Austral
    "Multiple states", # Various multi-state localities
    "Idaho, Nevada, Utah", # Sagebrush steppe across multiple states
    "California, Idaho, Nevada, Oregon, Utah", # Multiple states
    "Midi-Pyrénées", # Localities: Bagnères-de-Bigorre, Uzer
    "North Island", # Several specific localities
    "Pyrenees",    # Locality: Camurac
    "Rio Negro and Neuquen", # Nahuel Huapi and Lanín National Parks
    "Northern Cape/Western Cape", # Namaqualand region
    "Salta/Jujuy", # Río and Mango populations
    "São Paulo/Rio de Janeiro border",
    "Central Chile", # Multiple sites spanning regions
    "Kerala and Tamil Nadu",
    "Western Australia and Northern Territory",
    "Honshu",      # Mt. Izumigatake localities
    "Maharashtra and Goa",
    "Shannxi Province",
    "Rhône-Alpes", # Col du Lautaret locality
    "Miyagi Prefecture and Yamagata Prefecture",
    "Natal",       # Himeville Nature Reserve
    "Cape Province", # Multiple historical locations
    "Amami Islands",
    "Ryukyu Islands",
    "Yakushima", 
    "Kyushu",      # Kamae area/region
    "Crimean Peninsula",
    "Kerala and Karnataka",
    "IV Región",   # Multiple localities 
    "Hyogo, Okayama, and Tottori Prefectures",
    "Multiple prefectures",
    "Colchagua",   # Near Las Peñas
    "Talca",       # Natural Preserve Altos de Lircay
    "South Island", # Multiple specific locations
    "Ryukyu Archipelago", # Multiple islands
    "Qinghai-Tibet", # Alpine Meadow Ecosystem Research Station
    "Western Cape and Eastern Cape", # 11 populations across regions
    "Bonin Islands",
    "Kinki",       # Mt. Hyonosen area
    "Rio de Janeiro/Minas Gerais", # Itatiaia National Park
    "Eastern Cape and KwaZulu-Natal", # East coast sites
    "Washington and Oregon", # Columbia River Gorge
    "Multiple including Gauteng and KwaZulu-Natal",
    "Banks Peninsula", # Prices Valley
    "Victoria and New South Wales",
    "Chiloé",      # Multiple specific localities in northern Chiloé
    "VIII Region", # Cobquecura
    "Washington and Idaho",
    "Idaho and Washington", # Same area different naming
    "Oregon/Washington/Idaho", # Sagebrush-bunchgrass steppe
    "Western Ghats",
    "Sikkim and Darjeeling", # Multiple tea plantations
    "Maule and Los Ríos Regions",
    "Provence",    # Specific localities: Bagaud Island, Brégançon
    "Ogasawara Islands", # Anijima Island
    "Ryukyu archipelago", # Multiple islands
    "Multiple prefectures", # Akao site and Uruno site
    "Hyogo, Osaka, Nara", # Osaka-Kobe metropolitan area
    "Region V",    # Multiple localities all in Valparaíso
    "multiple states", # 24 populations across states
    "Languedoc",   # St. Bauzille, La Clause, Hortus Mountain
    "Xisha Islands", # Yongxing Island
    "Alabama, Georgia, South Carolina, North Carolina",
    "Cape Province", # Multiple historical sites
    "California/Oregon border", # Intermontane valleys
    "multiple states" # Great Basin and Columbia Plateau
  ),
  state_ISO_code = c(
    NA,            # Historical Cape Province region
    "CL-VS",       # All localities in Valparaíso Region
    "CL-RM",       # Part of Santiago Metropolitan Region
    "CL-VS",       # Valparaíso Region including islands
    "CL-VS",       # Part of Valparaíso Region
    NA,            # Historical region
    "CL-LR",       # Los Ríos Region
    NA,            # Multiple states
    NA,            # Multiple states
    NA,            # Multiple states
    NA,            # Historical region
    "NZ-NTL",      # North Island
    NA,            # Mountain range across regions
    NA,            # Two provinces
    NA,            # Two provinces
    NA,            # Two provinces
    NA,            # Border region
    NA,            # Multiple regions
    NA,            # Two states
    NA,            # Two territories
    NA,            # Region rather than prefecture
    NA,            # Two states
    "CN-SN",       # Shaanxi Province
    NA,            # Historical region
    NA,            # Two prefectures
    "ZA-NL",       # Now part of KwaZulu-Natal
    NA,            # Historical province
    NA,            # Island group
    NA,            # Island group
    NA,            # Island
    NA,            # Region rather than prefecture
    NA,            # Disputed territory
    NA,            # Two states
    "CL-CO",       # Coquimbo Region
    NA,            # Multiple prefectures
    NA,            # Multiple prefectures
    "CL-LI",       # O'Higgins Region
    "CL-ML",       # Maule Region
    "NZ-STL",      # South Island
    NA,            # Multiple islands
    NA,            # Cross-border region
    NA,            # Two provinces
    NA,            # Island group
    NA,            # Region rather than prefecture
    NA,            # Two states
    NA,            # Two provinces
    NA,            # Two states
    NA,            # Multiple provinces
    "NZ-CAN",      # Canterbury Region
    NA,            # Two states
    "CL-LL",       # Los Lagos Region
    "CL-BI",       # Biobío Region
    NA,            # Two states
    NA,            # Two states
    NA,            # Three states
    NA,            # Geographic region
    NA,            # Cross-border region
    NA,            # Two regions
    "FRA",      # Now part of Provence-Alpes-Côte d'Azur
    NA,            # Island group
    NA,            # Island group
    NA,            # Multiple prefectures
    NA,            # Multiple prefectures
    "CL-VS",       # Valparaíso Region
    NA,            # Multiple states
    NA,            # Historical region
    NA,            # Island group
    NA,            # Multiple states
    NA,            # Historical province
    NA,            # Border region
    NA             # Multiple states
  )
) %>%
  filter(!is.na(state_ISO_code))

```

The only hiccup left is that Hong Kong does not show up either in the countries or states map. Since we are interested in global patterns, I just joined the results with the closes Chinese province (Guandong) for visualization purposes.

Now let's combine both:
```{r}
state_iso_mapping = rbind(state_iso_mapping %>% select(state_province = original_state_name, state_ISO_code),
                          additional_state_mapping %>% select(state_province = original_state_name, state_ISO_code),
                          existing_states %>% select(state_province = state_name, state_ISO_code)) %>%
  distinct()
state_iso_mapping
```

Now let's add a column to combined data that will tell us which code we will use to map:
```{r}
filtered_state_level_records = combined_records %>% 
  left_join(state_iso_mapping, by = "state_province") %>%
  filter(!is.na(country_ISO_code), !((country_ISO_code %in% higher_lat_range_codes) & is.na(state_ISO_code))) %>%
  mutate(grouping_iso_code = ifelse(is.na(state_ISO_code),country_ISO_code,state_ISO_code))

filtered_state_level_records <- filtered_state_level_records %>%
  mutate(
    grouping_iso_code = case_when(
      # Check if the code starts with "FR-" but is not one of the overseas territories
      startsWith(grouping_iso_code, "FR-") & 
      !grouping_iso_code %in% c("FR-GF", "FR-MQ", "FR-GP", "FR-RE", "FR-YT") ~ "FRA",
      # Keep all other codes unchanged
      TRUE ~ grouping_iso_code
    )
  )

filtered_state_level_records
```

## Reconcile plant names

Let's now filter even further and delete records for which we cannot parse plant names. We will get plant names from the World Flora Online API first:

Accessory functions
```{r}
# Clean name function - removes HTML tags and formats the name correctly
clean_name <- function(name) {
  if (is.null(name) || length(name) == 0 || is.na(name)) return(NA)
  
  # Extract name and author parts
  name_part <- sub('.*<span class="wfo-name">(.*?)</span>.*', '\\1', name)
  author_part <- sub('.*<span class="wfo-name-authors"[^>]*>(.*?)</span>.*', '\\1', name)
  author_part <- if (author_part == name) "" else author_part
  
  # Clean up HTML tags
  name_part <- gsub("<i>([^<]+)</i>", "\\1", name_part)
  name_part <- gsub('<span class="wfo-name-rank">([^<]+)</span>', "\\1", name_part)
  name_part <- gsub("<[^>]+>", "", name_part)
  name_part <- gsub("\\s+", " ", name_part)
  name_part <- trimws(name_part)
  
  author_part <- gsub("<[^>]+>", "", author_part)
  author_part <- gsub("\\s+", " ", author_part)
  author_part <- trimws(author_part)
  
  if (nchar(author_part) > 0) {
    paste(name_part, author_part)
  } else {
    name_part
  }
}

# Extract family name from path string
extract_family_from_path <- function(path_string) {
  if (is.null(path_string) || is.na(path_string)) return(NA)
  
  parts <- strsplit(path_string, "/")[[1]]
  family_names <- parts[grep("aceae$", parts)]
  if (length(family_names) > 0) {
    return(family_names[1])
  }
  return(NA)
}

# Extract family information from name data
extract_family <- function(name_data) {
  # Handle cases with NA currentPreferredUsage first
  if (is.data.frame(name_data) && "currentPreferredUsage" %in% names(name_data) && 
      all(is.na(name_data$currentPreferredUsage))) {
    return(list(name = NA, id = NA))
  }
  
  # Handle unplaced names
  if (!is.null(name_data$role) && name_data$role == "unplaced") {
    if (!is.null(name_data$hasAssociatedGenusName)) {
      # For data frame structure
      if (is.data.frame(name_data$hasAssociatedGenusName)) {
        path_string <- name_data$hasAssociatedGenusName$currentPreferredUsage$pathString[1]
        if (!is.null(path_string) && !is.na(path_string)) {
          family_name <- extract_family_from_path(path_string)
          if (!is.na(family_name)) {
            return(list(name = family_name, id = NA))
          }
        }
      } 
      # For list structure
      else if (is.list(name_data$hasAssociatedGenusName)) {
        if (!is.null(name_data$hasAssociatedGenusName$currentPreferredUsage$pathString)) {
          family_name <- extract_family_from_path(name_data$hasAssociatedGenusName$currentPreferredUsage$pathString)
          if (!is.na(family_name)) {
            return(list(name = family_name, id = NA))
          }
        }
      }
    }
    return(list(name = NA, id = NA))
  }
  
  # Handle placed names
  if (!is.null(name_data$currentPreferredUsage)) {
    # Try path string first
    if (!is.null(name_data$currentPreferredUsage$pathString)) {
      family_name <- extract_family_from_path(name_data$currentPreferredUsage$pathString)
      if (!is.na(family_name)) {
        # Try to find family ID by traversing up the hierarchy
        current <- name_data$currentPreferredUsage$isPartOf
        while (!is.null(current)) {
          if (!is.null(current$hasName)) {
            name_html <- current$hasName$fullNameStringHtml
            name <- sub('.*<span class="wfo-name">([^<]+)</span>.*', '\\1', name_html)
            name <- trimws(name)
            
            if (grepl("aceae$", name)) {
              return(list(name = name, id = current$hasName$id))
            }
          }
          current <- current$isPartOf
        }
        return(list(name = family_name, id = NA))
      }
    }
  }
  
  return(list(name = NA, id = NA))
}

# Make API request with retry logic
make_api_request <- function(query, retries_left = 3, retry_delay = 2) {
  tryCatch({
    response <- POST(
      url = "https://list.worldfloraonline.org/gql.php",
      body = list(query = query),
      encode = "json",
      add_headers(
        "Content-Type" = "application/json",
        "Accept" = "application/json"
      ),
      timeout(20)
    )
    
    if (http_error(response)) {
      stop(sprintf("HTTP error: %s", http_status(response)$message))
    }
    
    # Check for GraphQL errors in response
    content <- fromJSON(rawToChar(response$content))
    if (!is.null(content$errors)) {
      stop(sprintf("GraphQL errors: %s", 
                  paste(sapply(content$errors, function(e) e$message), collapse="; ")))
    }
    
    return(response)
    
  }, error = function(e) {
    if (retries_left > 0) {
      warning(sprintf("Request failed, retrying in %d seconds. Retries left: %d. Error: %s", 
                     retry_delay, retries_left, e$message))
      Sys.sleep(retry_delay)
      return(make_api_request(query, retries_left - 1))
    } else {
      stop(sprintf("All retries exhausted. Last error: %s", e$message))
    }
  })
}

# Complete family IDs for results
complete_family_ids <- function(df) {
  # First pass: use known family IDs
  known_families <- df[!is.na(df$family_id), c("family", "family_id")]
  family_id_map <- unique(known_families)
  
  for (i in 1:nrow(df)) {
    if (is.na(df$family_id[i]) && !is.na(df$family[i])) {
      matching_row <- family_id_map[family_id_map$family == df$family[i], ]
      if (nrow(matching_row) > 0) {
        df$family_id[i] <- matching_row$family_id[1]
      }
    }
  }
  
  # Second pass: lookup remaining families
  remaining_families <- unique(df$family[is.na(df$family_id) & !is.na(df$family)])
  
  if (length(remaining_families) > 0) {
    queries <- sapply(seq_along(remaining_families), function(i) {
      sprintf('
        f%d: taxonNameSuggestion(termsString: "%s", limit: 1) {
          id
          fullNameStringHtml
          role
        }
      ', i, remaining_families[i])
    })
    
    query <- sprintf('{ %s }', paste(queries, collapse = "\n"))
    response <- make_api_request(query)
    response_data <- fromJSON(rawToChar(response$content))
    
    for (i in seq_along(remaining_families)) {
      family_name <- remaining_families[i]
      result <- response_data$data[[paste0("f", i)]]
      if (length(result) > 0) {
        df$family_id[df$family == family_name & is.na(df$family_id)] <- result$id[1]
      }
    }
  }
  
  return(df)
}

# Create batch query for name matching
create_match_query <- function(batch_names) {
  queries <- sapply(seq_along(batch_names), function(i) {
    sprintf('
      n%d: taxonNameMatch(
        inputString: "%s"
        fuzzyNameParts: 3
        fuzzyAuthors: 3
        checkHomonyms: true
      ) {
        match {
          id
          fullNameStringHtml
          role
          currentPreferredUsage {
            id
            pathString
            hasName {
              id
              fullNameStringHtml
            }
            isPartOf {
              hasName {
                id
                fullNameStringHtml
              }
              isPartOf {
                hasName {
                  id
                  fullNameStringHtml
                }
              }
            }
          }
          hasAssociatedGenusName {
            id
            fullNameStringHtml
            currentPreferredUsage {
              pathString
              hasName {
                id 
                fullNameStringHtml
              }
            }
          }
        }
        candidates {
          id
          fullNameStringHtml
          role
          currentPreferredUsage {
            id
            pathString
            hasName {
              id
              fullNameStringHtml
            }
            isPartOf {
              hasName {
                id
                fullNameStringHtml
              }
              isPartOf {
                hasName {
                  id
                  fullNameStringHtml
                }
              }
            }
          }
        }
      }
    ', i, gsub('"', '\\\\"', batch_names[i]))  # Escape quotes in plant names
  })
  
  sprintf('{ %s }', paste(queries, collapse = "\n"))
}
```


Main code to query WFO:
```{r}
# Main function for checking plant names against WFO database
check_plant_names <- function(plant_names, batch_size = 10, max_retries = 3, retry_delay = 2) {
  # Required packages
  if (!require("httr")) install.packages("httr")
  if (!require("jsonlite")) install.packages("jsonlite")
  library(httr)
  library(jsonlite)
  
  # Split names into batches
  batches <- split(plant_names, ceiling(seq_along(plant_names) / batch_size))
  
  # Process batches and combine results
  all_results <- list()
  
  for (batch_num in seq_along(batches)) {
    current_batch <- batches[[batch_num]]
    message(sprintf("Processing batch %d of %d...", batch_num, length(batches)))
    
    # Try to process the batch
    batch_results <- tryCatch({
      # Create and execute query
      query <- create_match_query(current_batch)
      response <- make_api_request(query, max_retries, retry_delay)
      response_data <- fromJSON(rawToChar(response$content))
      
      # Process each name in the batch
      lapply(names(response_data$data), function(node_name) {
        match_data <- response_data$data[[node_name]]
        original_query <- current_batch[as.numeric(sub("n", "", node_name))]
        process_name_result(match_data, original_query)
      })
    }, error = function(e) {
      message(sprintf("Batch failed, processing names individually..."))
      
      # Process each name individually
      lapply(seq_along(current_batch), function(i) {
        name <- current_batch[i]
        tryCatch({
          query <- create_match_query(name)
          response <- make_api_request(query, max_retries, retry_delay)
          response_data <- fromJSON(rawToChar(response$content))
          match_data <- response_data$data$n1
          process_name_result(match_data, name)
        }, error = function(e) {
          # On error, return NA for all fields
          list(
            original_query = name,
            wfo_id = NA,
            is_synonym = NA,
            original_name = NA,
            accepted_name = NA,
            accepted_id = NA,
            family = NA,
            family_id = NA
          )
        })
      })
    })
    
    all_results <- c(all_results, batch_results)
  }
  
  # Convert results to data frame
  result_df <- do.call(rbind, lapply(all_results, function(x) {
    data.frame(
      original_query = x$original_query,
      wfo_id = x$wfo_id,
      is_synonym = x$is_synonym,
      original_name = x$original_name,
      accepted_name = x$accepted_name,
      accepted_id = x$accepted_id,
      family = x$family,
      family_id = x$family_id,
      stringsAsFactors = FALSE
    )
  }))
  
  # Complete missing family IDs
  result_df <- complete_family_ids(result_df)
  
  # Post-processing: propagate family information within genera
  # Extract genus from original_name (first word)
  result_df$genus <- sapply(strsplit(result_df$original_name, " "), function(x) if(length(x) > 0) x[1] else NA)
  
  # Create a genus-to-family mapping from known entries
  genus_family_map <- result_df[!is.na(result_df$family) & !is.na(result_df$genus), 
                               c("genus", "family", "family_id")] %>%
    unique()
  
  # Update missing family information based on genus
  for(i in 1:nrow(result_df)) {
    if(is.na(result_df$family[i]) && !is.na(result_df$genus[i])) {
      matching_row <- genus_family_map[genus_family_map$genus == result_df$genus[i], ]
      if(nrow(matching_row) > 0) {
        result_df$family[i] <- matching_row$family[1]
        result_df$family_id[i] <- matching_row$family_id[1]
      }
    }
  }
  
  # Remove the temporary genus column
  result_df$genus <- NULL
  
  return(result_df)
}

# Helper function to process a single name result
process_name_result <- function(match_data, original_query) {
  # Get the best match - either exact match or first candidate
  name_data <- if (!is.null(match_data$match)) {
    match_data$match
  } else if (is.data.frame(match_data$candidates) && nrow(match_data$candidates) > 0) {
    # For data frame candidates, handle NA currentPreferredUsage
    candidates <- match_data$candidates
    if (all(is.na(candidates$currentPreferredUsage))) {
      # Take first unplaced candidate
      candidates[1,]
    } else {
      # Prefer accepted names that have currentPreferredUsage
      valid_idx <- !is.na(candidates$currentPreferredUsage)
      if (any(valid_idx)) {
        valid_candidates <- candidates[valid_idx,]
        accepted_idx <- which(valid_candidates$role == "accepted")
        if (length(accepted_idx) > 0) {
          valid_candidates[accepted_idx[1],]
        } else {
          valid_candidates[1,]
        }
      } else {
        candidates[1,]
      }
    }
  } else {
    return(list(
      original_query = original_query,
      wfo_id = NA,
      is_synonym = NA,
      original_name = NA,
      accepted_name = NA,
      accepted_id = NA,
      family = NA,
      family_id = NA
    ))
  }
  
  # Extract family information
  family_info <- extract_family(name_data)
  
  # Handle unplaced names, NULL or NA currentPreferredUsage
  if (is.null(name_data$currentPreferredUsage) || 
      (is.atomic(name_data$currentPreferredUsage) && is.na(name_data$currentPreferredUsage)) ||
      (is.data.frame(name_data$currentPreferredUsage) && nrow(name_data$currentPreferredUsage) == 0) ||
      (is.logical(name_data$currentPreferredUsage) && all(is.na(name_data$currentPreferredUsage)))) {
    return(list(
      original_query = original_query,
      wfo_id = name_data$id,
      is_synonym = NA,
      original_name = clean_name(name_data$fullNameStringHtml),
      accepted_name = "Unplaced name",
      accepted_id = NA,
      family = family_info$name,
      family_id = family_info$id
    ))
  }
  
  # Determine if name is a synonym - but first check if hasName exists
  if (!is.null(name_data$currentPreferredUsage$hasName)) {
    is_synonym <- name_data$id != name_data$currentPreferredUsage$hasName$id
  } else {
    is_synonym <- NA
  }
  
  # Return processed result
  list(
    original_query = original_query,
    wfo_id = name_data$id,
    is_synonym = is_synonym,
    original_name = clean_name(name_data$fullNameStringHtml),
    accepted_name = clean_name(name_data$currentPreferredUsage$hasName$fullNameStringHtml),
    accepted_id = name_data$currentPreferredUsage$hasName$id,
    family = family_info$name,
    family_id = family_info$id
  )
}
```

Now we can apply this function. We will save the result when re-running this notebook to avoid unnecessary API calls.
```{r}

#plant_names = check_plant_names((filtered_state_level_records$plant_species %>% unique()))
#save(plant_names,file = "tables/WFO_results.Rdata")
load("tables/WFO_results.Rdata")
plant_names
```

Were there records that could not be retrieved?
```{r}
plant_names %>%
  filter(is.na(family))
```

Great, there are only a few cases, and the majority should be removed anyway since they are not single plant species.

```{r}
processed_plant_names = plant_names %>%
  filter(!is.na(family)) %>%
  select(plant_species = original_query,
         plant_species_original_wfo_id = wfo_id,
         plant_species_is_synonym = is_synonym,
         plant_speices_original_match = original_name,
         plant_species_current_name = accepted_name,
         plant_species_current_wfo_id = accepted_id,
         plant_family_name = family,
         plant_family_wfo_id = family_id
         )

filtered_state_level_records = filtered_state_level_records %>%
  left_join(processed_plant_names, by = 'plant_species') %>%
  filter(!is.na(family))

filtered_state_level_records
```
# Make plots

Now we will use this filtered dataset to make plots showing the global distribution of beetle flower visitors and pollinators, as well as the effect of sampling method and time.

## Plot maps

Now let's summarize records for each geographical group:
```{r}
iso_level_summaries = filtered_state_level_records %>%
  group_by(grouping_iso_code) %>%
  summarise(n_records=n(), beetle_fraction=mean(beetle_visitors,na.rm = T), beetle_poll_fraction = mean(beetle_pollinators,na.rm = T))

iso_level_summaries
```

Let's now combine this with geographical data.
```{r}
combined_geo_data <- bind_rows(
  ne_countries(scale = "medium", returnclass = "sf") %>%
    select(adm0_a3, name_long, geometry, label_x, label_y) %>%
    rename(grouping_iso_code = adm0_a3, geo_name = name_long) %>%
    mutate(geo_level = "country"),
  ne_states(returnclass = "sf") %>%
    select(iso_3166_2, name, geometry, label_x = longitude, label_y = latitude) %>%
    rename(grouping_iso_code = iso_3166_2, geo_name = name) %>%
    mutate(geo_level = "state")
)

combined_geo_data %>% as_tibble() 
```

Now let's combine both
```{r}
map_plot_data = left_join(iso_level_summaries, combined_geo_data)
map_plot_data
```

Now, let's plot our map for frequency of beetle flower visitors.
```{r}
world_map = world_sf %>%
  mutate(has_data = adm0_a3 %in% map_plot_data$grouping_iso_code) %>%
  select(has_data)

states_map = ne_states() %>%
  filter(adm0_a3 %in% higher_lat_range_codes[-1] | iso_3166_2 %in% c("FR-GF", "FR-MQ", "FR-GP", "FR-RE", "FR-YT")) %>%
  mutate(has_data = iso_3166_2 %in% map_plot_data$grouping_iso_code) %>%
  select(has_data)
  

p1 = ggplot() +
  # Base world map
  geom_sf(data = world_map, aes(fill = has_data), color = gray(0.5)) +
  geom_sf(data = states_map, aes(fill = has_data, alpha = has_data), color = gray(0.5)) +
  scale_alpha_manual(values = c("TRUE" = 1 , "FALSE" = 0)) +
  scale_fill_manual(values = c("TRUE" = gray(0.95), "FALSE" = gray(0.95))) +
  # Add bubbles at label positions
  geom_point(data = map_plot_data %>% filter(!is.na(n_records)),
             aes(x = label_x, y = label_y,
                 size = n_records, # use sqrt for area proportional scaling
                 color = beetle_fraction),
             alpha = 0.7) +
  # Customize appearance
  scale_size_continuous(name = "Number of Records",
                       breaks = c(1, 10, 100),
                       labels = c(1, 10, 100),
                       range = c(0, 3.5)) +
  scale_color_viridis_c(name = "Proportion of records",
                        limits = c(0, 1),
                        breaks = c(0,0.5,1),
                        labels = c(0,0.5,1),
                        option = "turbo") +
  # Theme customization
  labs(title = "Plants with beetle flower visitors", x = NULL, y = NULL) +
  theme_minimal() +
  guides(fill = "none", alpha="none",
         size = guide_legend(
    title.position = "top",
    title.hjust = 0.5  # Centers the title horizontally
  ),
  color = guide_colorbar(
    title.position = "top",
    title.hjust = 0.5,
    barwidth = 4,
    barheight = 0.5
  )) +
  theme(text = element_text(size=8),
        legend.position = "bottom",
    legend.box = "horizontal",
    legend.key.size = unit(0.4, "lines"),
    legend.spacing.x = unit(0.1, "cm"),
    legend.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt")) +
  ylim(-60,90)

p1
```
Now the same bu for pollinators
```{r}

p2 = ggplot() +
  # Base world map
  geom_sf(data = world_map, aes(fill = has_data), color = gray(0.5)) +
  geom_sf(data = states_map, aes(fill = has_data, alpha = has_data), color = gray(0.5)) +
  scale_alpha_manual(values = c("TRUE" = 1 , "FALSE" = 0)) +
  scale_fill_manual(values = c("TRUE" = gray(0.95), "FALSE" = gray(0.95))) +
  # Add bubbles at label positions
  geom_point(data = map_plot_data %>% filter(!is.na(n_records)),
             aes(x = label_x, y = label_y,
                 size = n_records, # use sqrt for area proportional scaling
                 color = beetle_poll_fraction),
             alpha = 0.7) +
  # Customize appearance
  scale_size_continuous(name = "Number of Records",
                       range = c(0, 3.5)) +
  scale_color_viridis_c(name = "Proportion of records",
                        limits = c(0, 1),
                        option = "turbo") +
  # Theme customization
  labs(title = "Plants with beetle pollinators", x = NULL, y = NULL) +
  theme_minimal() +
  guides(fill = "none", alpha="none", color = "none", size = "none") +
  theme(legend.position = 'bottom', text = element_text(size=8)) +
  ylim(-60,90)

p2
```

```{r}
# Now combine the plots with specific arrangement of legends
combined_map <- patchwork::wrap_plots(
  p1,
  p2,
  ncol = 1
) +
  plot_layout(guides = 'collect') +  # This collects and combines similar legends
  plot_annotation(
    theme = theme(
      legend.position = 'bottom',
    )
  )

combined_map
```

## Plot dataset characteristics

Now let's show other aspects of the dataset. How many studies are included in this figure?
```{r}
filtered_state_level_records %>% pull(bibtex_id) %>% unique() %>% length()
```

And how many records?
```{r}
nrow(filtered_state_level_records)
```


What is the distribution of number of records per study?
```{r}
filtered_state_level_records %>% group_by(bibtex_id) %>%
  summarise(N=n()) %>%
  ggplot() + geom_histogram(aes(x=N))
```

What is the proportion of records with unbiased sampling, split between those that have found beetles and those that have not?
```{r}
bias_vs_beetle = filtered_state_level_records %>%
  select(beetle_visitors, methods_unbiased) %>%
  group_by_all() %>%
  summarise(N=n()) %>%
  group_by(methods_unbiased) %>%
  mutate(p = N/sum(N)) %>%
  arrange(methods_unbiased)

bias_vs_beetle
```

```{r}
bias_vs_beetlepollinator = filtered_state_level_records %>%
  select(beetle_pollinators, methods_unbiased) %>%
  group_by_all() %>%
  summarise(N=n()) %>%
  group_by(methods_unbiased) %>%
  mutate(p = N/sum(N)) %>%
  arrange(methods_unbiased)

bias_vs_beetlepollinator
```
Now let's plot those tables as barplots

```{r}
barplot_df = filtered_state_level_records %>%
  select(beetle_visitors, beetle_pollinators, methods_unbiased) %>%
  group_by_all() %>%
  summarise(N = n(), .groups = 'drop') %>%  # Added .groups = 'drop' to avoid grouping warning
  group_by(methods_unbiased) %>%
  mutate(p = N/sum(N)) %>%
  arrange(methods_unbiased) %>%
  mutate(beetle_status = case_when(
    beetle_pollinators ~ "beetle pollinator",
    beetle_visitors ~ "beetle nonpollinator",
    TRUE ~ "no beetles"
  )) %>%
  ungroup() %>%
  mutate(
    # Create a new factor with updated labels while preserving the correct mapping
    beetle_status = factor(
      case_when(
        beetle_status == "no beetles" ~ "none",
        beetle_status == "beetle nonpollinator" ~ "nonpollinators",
        beetle_status == "beetle pollinator" ~ "pollinators"
      ),
      # Set the order we want the categories to appear in the stack
      levels = c("pollinators", "nonpollinators", "none")
    )
  )

p_bars = ggplot(barplot_df, aes(x = methods_unbiased, y = N, fill = beetle_status)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(
    aes(label = scales::percent(p, accuracy = 1)),
    position = position_stack(vjust = 0.5),
    color = "white",
    size = 2
  ) +
  scale_fill_brewer(palette = "Set2") +
  scale_x_discrete(
    labels = c("FALSE" = "targeted", 
              "TRUE" = "broad")
  ) +
  labs(
    x = "Flower Visitor Sampling Method",
    y = "Number of records in literature",
    fill = "Beetle flower visitors"
  ) +
  theme_minimal() +
  guides(fill = guide_legend(title.position = "top",
    title.hjust = 0.5)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        text = element_text(size=8),
               legend.position = "bottom",
    legend.box = "horizontal",
    legend.key.size = unit(0.4, "lines"),
    legend.spacing.x = unit(0.1, "cm"),
    legend.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt")
        )

p_bars
```

Finally, let's have a look at the years of publication in the studies included. Let's first load bibliographic information for all studies (including those we failed to obtain data). Since R parsers do not seem to be able to read this bibtex appropriately, let's use python for that, and then communicate with R through reticulate.


```{python}
from pybtex.database.input import bibtex
import pandas as pd

def parse_bibtex(file_path):
    parser = bibtex.Parser()
    bib_data = parser.parse_file(file_path)
    
    # Convert to a more R-friendly format
    entries = []
    for key, entry in bib_data.entries.items():
        authors = " and ".join(str(person) for person in entry.persons.get('author', []))
        entries.append({
            'key': key,
            'authors': authors,
            'year': entry.fields.get('year', ''),
            'title': entry.fields.get('title', ''),
            'journal': entry.fields.get('journal', '')
        })
    
    return pd.DataFrame(entries)

# Parse your BibTeX file
df = parse_bibtex("pdfs/export_20250131/export_20250131.updated.bib")
```

```{r}
bibliography = as_tibble(py$df) %>%
  mutate(year = as.numeric(year)) %>%
  select(bibtex_id = key,year_published = year)
bibliography
```
Now let's join this bibliographic information with the dataset containing records:
```{r}
year_plot_df <- filtered_state_level_records %>% 
  # Join with bibliography data
  left_join(bibliography) %>%
  # grouping to count papers per year and beetle status
  group_by(year_published, beetle_visitors, beetle_pollinators) %>%
  summarise(
    N = n(),
    .groups = "drop"  # Explicitly drop grouping again
  ) %>%
  # Create the beetle status factor
  mutate(beetle_status = factor(case_when(
    beetle_pollinators ~ "pollinators",
    beetle_visitors ~ "nonpollinators",
    TRUE ~ "none"
  ),
  levels = c("pollinators", "nonpollinators", "none")
  )) %>%
  filter(year_published < 2025)

year_plot_df
```

Let's visualize the number of records over time.

```{r}
p_years = ggplot(year_plot_df, aes(x = year_published, y = N, fill = beetle_status)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_brewer(palette = "Set2") +
  scale_x_continuous(
    breaks = seq(min(year_plot_df$year_published), 
                max(year_plot_df$year_published), 
                by = 10)  # Creates breaks every 10 years
  ) +
  labs(
    x = "Year of publication",
    y = "Number of records in literature",
    fill = "Beetle flower visitors"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.minor = element_blank(),  # Removes minor grid lines for cleaner look
    legend.position = "bottom",
    text = element_text(size = 8)
  ) +
  scale_x_continuous(breaks=seq(1930,2020,by=10))

p_years
```
Now let's visualize the proportion of records of each kind over time.

```{r}
# Calculate proportions within each year
year_plot_df_prop <- year_plot_df %>%
  # Group by year to calculate proportions
  group_by(year_published) %>%
  # Calculate the proportion for each beetle status within each year
  mutate(
    proportion = N / sum(N)
  ) %>%
  ungroup()

ggplot(year_plot_df_prop, aes(x = year_published, y = proportion, fill = beetle_status)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_brewer(palette = "Set2") +
  scale_x_continuous(
    breaks = seq(min(year_plot_df_prop$year_published), 
                max(year_plot_df_prop$year_published), 
                by = 10)
  ) +
  # Add a percentage scale for the y-axis
  scale_y_continuous(
    labels = scales::percent,  # Format y-axis labels as percentages
    breaks = seq(0, 1, 0.2)    # Set breaks at 20% intervals
  ) +
  labs(
    x = "Year of publication",
    y = "Proportion of records",
    fill = "Beetle flower visitors"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.minor = element_blank()
  )

```


Let's now join the two plots

```{r}
# Now combine the plots with specific arrangement of legends
combined_bars <- patchwork::wrap_plots(
  p_bars,
  p_years + guides(fill='none'),
  ncol = 1
) +
  plot_layout(guides = 'collect') +  # This collects and combines similar legends
  plot_annotation(
    theme = theme(
      legend.position = 'bottom',  # Put all legends on the right
    )
  )

combined_bars
```
## Save plots

Finally, let's save all plots to use as figures in the paper.

```{r}
combined_plot = cowplot::plot_grid(combined_map, combined_bars, ncol =2,rel_widths = c(2,1.5))
combined_plot
```

```{r}
ggsave(filename = 'figures/fig1.pdf',device = 'pdf',width = 6.33,units = 'in', height = 5, useDingbats=F)
ggsave(combined_map, filename = 'figures/fig1A.pdf',device = 'pdf',width = 4,units = 'in', height = 6, useDingbats=F)
ggsave(combined_bars, filename = 'figures/fig1BC.pdf',device = 'pdf',units = 'mm', height = 118, width=59, useDingbats=F)
```

# Produce supplementary table 1

Here we will summarize information by beetle family, to help us produce Supplementary Table 1 by identifying records of beetle flower visitors that have not yet been properly recognized as such in the literature.

Let's now list all of the beetle families mentioned across the studies.

```{r}
filtered_state_level_records$beetle_families %>%
  unlist %>%
  unique %>%
  sort %>%
  paste(collapse="\n") %>%
  cat()
```

After filtering the list manually, the following do not seem to have been recorded as flower visitors before in the literature:
```{r}
target_families <- c("Bostrichidae", "Byrrhidae", "Byturidae", "Cicindelidae",
                     "Cucujidae", "Dascillidae", "Dytiscidae",
                    "Hybosoridae", "Latridiidae", "Lathridiidae", "Leiodidae", "Lymexylidae",
                    "Monotomidae",  "Phalacradidae", "Phalacridae",
                    "Phalacrididae", "Silvanidae", "Silvannidae","Ptiliidae")

target_families
```

Let's find which studies mention them:
```{r}
newly_reported_families = filtered_state_level_records %>%
  mutate(
    # Check if any family matches (returns TRUE/FALSE)
    has_target_family = map_lgl(beetle_families, 
                               ~any(.x %in% target_families)),
    
    # Get the actual matching families as a string
    matching_families = map_chr(beetle_families, function(x) {
      matches <- x[x %in% target_families]
      if(length(matches) > 0) str_c(matches, collapse = ", ") else NA_character_
    })
  ) %>%
  filter(has_target_family) %>%
  mutate(families_string = map_chr(matching_families, ~str_c(.x, collapse = ", "))) %>%
  select(bibtex_id,families_string) %>%
  distinct() %>%
  arrange(families_string)

newly_reported_families
```

Let's now print a report of these studies including any noteworthy facts the LLM was able to pull:

```{r}
format_reference <- function(bibtex_id, df) {
  # Get bibliographic record
  record <- df %>% 
    filter(key == bibtex_id) %>%
    as.list()
  
  # Get beetle families for this specific reference
  beetle_fams <- filtered_state_level_records %>%
    filter(bibtex_id == !!bibtex_id) %>%
    pull(beetle_families) %>%
    unlist() %>%
    unique() %>%
    paste(collapse = ", ")
  
  # Get noteworthy fact
  fact <- review_data[[bibtex_id]]$json$noteworthy_beetle_fact
  
  # Format author names more nicely
  authors <- str_split(record$authors, " and ")[[1]]
  
  # Create formatted string
  cat(sprintf("Reference ID: %s\n", record$key),
      sprintf("Authors: %s\n", paste(authors, collapse = "; ")),
      sprintf("Year: %s\n", record$year),
      sprintf("Title: %s\n", record$title),
      sprintf("Journal: %s\n", record$journal),
      sprintf("Beetle families: %s\n", beetle_fams),
      if(!is.null(fact)) sprintf("Noteworthy fact: %s\n", fact),
      "\n",
      sep = "")
}

# Function to print all references
print_references <- function(bibtex_ids, df) {
  walk(bibtex_ids, ~format_reference(.x, df))
}

# Usage:
newly_reported_families %>%
  mutate(bibtex_id = as.character(bibtex_id)) %>%
  arrange(tolower(bibtex_id)) %>%
  pull(bibtex_id) %>%
  print_references(py$df)
```



# Supplementary Table 2

To assist producing supplementary table 2, we will now summarize the data by plant family.

Finally, let's list all plant families in our dataset, all families with beetle flower visitors and all families with beetle pollinators. Let's start by grouping references for each case:

```{r}
references_for_plant_families <- filtered_state_level_records %>%
  select(plant_family = plant_family_name, beetle_visitors, beetle_pollinators, bibtex_id, beetle_families) %>%
  group_by(plant_family, beetle_visitors, beetle_pollinators) %>%
  summarise(
    bibtex_ids = list(unique(bibtex_id)),
    beetle_families = list(unique(unlist(beetle_families[!sapply(beetle_families, is.null)]))),
    .groups = 'drop'
  )

references_for_plant_families
```

We will now reconcile beetle family names  following Cai et al. Let's create a function to map names read by Sonnet to the accepted names. For the cases in which the beetle is not a family, but a lower-level taxon, we will map to the corresponding family.

```{r}
# Define the mapping of various names to their accepted families
beetle_names <- tibble(
  original_name = c(
    "Acmaeoderidae", "Alleculidae", "Alleculiidae", "Allecullidae", "Allecuninae",
    "Allocorynidae", "Alticidae", "Amecocerus spp.", "Anaspidae", "Anobiidae",
    "Anthicidae", "Anthribidae", "Attelabidae", "Belidae", "Boganiidae",
    "Bostrichidae", "Brenthidae", "Brentidae", "Bruchidae", "Bruchinae",
    "Buprestidae", "Buprestridae", "Byrrhidae", "Byturidae", "Cantharidae",
    "Carabidae", "Carpophyllus", "Carpophylus", "Cebrionidae", "Cephaloidae",
    "Cerambicidae", "Cerambicydae", "Cerambycidae", "Cetonidae", "Cetoniidae",
    "Cetoniinae", "Chrysomelidae", "Cicindelidae", "Cleridae", "Cleroidea",
    "Coccidellidae", "Coccinelidae", "Coccinellidae", "Colegtera", "Coleoptera",
    "Colydidae", "Colydiidae", "Corylophidae", "Crioceridae", "Cryptophagidae",
    "Cryptophilidae", "Crysomelidae", "Cucujidae", "Cucurlionidae", "Culculionidae",
    "Curculionidae", "Curculionoidea", "Dascillidae", "Dasytaeidae", "Dasytidae",
    "Dermestidae", "Dryophthoridae", "Dynastidae", "Dynastinae", "Dytiscidae",
    "Elateridae", "Elatyridae", "Erotylidae", "Eucmenidae", "Eucnemanidae",
    "Galerucidae", "Galerucinae", "Geotrupidae", "Glaphyridae", "Halodidae",
    "Helodidae", "Histeridae", "Hopliidae", "Hybosoridae", "Hydrophilidae",
    "Kateretidae", "Lagridae", "Lagriidae", "Lampyridae", "Languriidae",
    "Lathridiidae", "Latridiidae", "Leiodidae", "Lycidae", "Lyctidae",
    "Lymexylidae", "Malachidae", "Malachiidae", "Malachiinae", "Maloidea",
    "Megalopodidae", "Melachiidae", "Melandryidae", "Meliodae", "Meliridae",
    "Meloidae", "Meloididae", "Melolinthinae", "Melolonthidae", "Melolontidae",
    "Melotonthidae", "Melyridae", "Monotomidae", "Mordelidae", "Mordellidae",
    "Mycetophagidae", "Mycteridae", "Mylabridae", "Nitidulidae", "none",
    "not specified", "Oedemeridae", "Oedemeriidae", "Oedomeridae", "Oemideridae",
    "Orsodacnidae", "Pentatomidae", "Phalacradidae", "Phalacridae", "Phalacrididae",
    "Ptiliidae", "Ptilliidae", "Ptinidae", "Pythidae", "Rhipiphoridae",
    "Rhizophagidae", "Ripiphoridae", "Ripriphoridae", "Rutelidae", "Rutelinae",
    "Scarabaeidae", "Scarabaeidae (Rutelinae)", "Scarabeidae", "Scarabidae",
    "Scaraboidea", "Scirtidae", "Scolytidae", "Scraptiidae", "Scydmaenidae",
    "Sessinia", "Silphidae", "Silvanidae", "Silvannidae", "Staphilinidae",
    "Staphylinidae", "Sylphidae", "Sylvaniidae", "Tenebrionidae", "Trogossitidae",
    "Undetermined", "Unidentified Coleoptera", "unspecified", "Zopheridae"
  ),
  accepted_family = c(
    "Buprestidae",           # Acmaeoderidae -> now part of Buprestidae
    "Tenebrionidae",         # Alleculidae -> now Tenebrionidae (Alleculinae)
    "Tenebrionidae",         # Alleculiidae -> misspelling
    "Tenebrionidae",         # Allecullidae -> misspelling
    "Tenebrionidae",         # Allecuninae -> misspelling
    "Belidae",                      # Allocorynidae -> Belidae
    "Chrysomelidae",         # Alticidae -> now part of Chrysomelidae
    "Melyridae",                      # Amecocerus spp. -> Melyridae
    "Scraptiidae",           # Anaspidae -> subfamily Anaspidinae of Scraptiidae
    "Ptinidae",             # Anobiidae -> now Ptinidae
    "Anthicidae",           # Anthicidae
    "Anthribidae",          # Anthribidae
    "Attelabidae",          # Attelabidae
    "Belidae",              # Belidae
    "Boganiidae",           # Boganiidae
    "Bostrichidae",         # Bostrichidae
    "Brentidae",            # Brenthidae -> now Brentidae
    "Brentidae",            # Brentidae
    "Chrysomelidae",        # Bruchidae -> now subfamily of Chrysomelidae
    "Chrysomelidae",        # Bruchinae -> subfamily of Chrysomelidae
    "Buprestidae",          # Buprestidae
    "Buprestidae",          # Buprestridae -> misspelling
    "Byrrhidae",            # Byrrhidae
    "Byturidae",            # Byturidae
    "Cantharidae",          # Cantharidae
    "Carabidae",            # Carabidae
    "Nitidulidae",          # Carpophyllus -> misspelling of Carpophilus (Nitidulidae)
    "Nitidulidae",          # Carpophylus -> misspelling of Carpophilus (Nitidulidae)
    "Elateridae",           # Cebrionidae -> now part of Elateridae
    "Stenotrachelidae",     # Cephaloidae -> subfamily of Stenotrachelidae
    "Cerambycidae",         # Cerambicidae -> misspelling
    "Cerambycidae",         # Cerambicydae -> misspelling
    "Cerambycidae",         # Cerambycidae
    "Scarabaeidae",         # Cetonidae -> now part of Scarabaeidae
    "Scarabaeidae",         # Cetoniidae -> now part of Scarabaeidae
    "Scarabaeidae",         # Cetoniinae -> subfamily of Scarabaeidae
    "Chrysomelidae",        # Chrysomelidae
    "Cicindelidae",            # Cicindelidae -> now part of Carabidae
    "Cleridae",             # Cleridae
    NA,                     # Cleroidea -> superfamily, not family
    "Coccinellidae",        # Coccidellidae -> misspelling
    "Coccinellidae",        # Coccinelidae -> misspelling
    "Coccinellidae",        # Coccinellidae
    NA,                     # Colegtera -> misspelling of order name
    NA,                     # Coleoptera -> order name, not family
    "Zopheridae",           # Colydidae -> now part of Zopheridae
    "Zopheridae",           # Colydiidae -> now part of Zopheridae
    "Corylophidae",         # Corylophidae
    "Chrysomelidae",        # Crioceridae -> now subfamily of Chrysomelidae
    "Cryptophagidae",       # Cryptophagidae
    "Erotylidae",           # Cryptophilidae -> subfamily of Erotylidae
    "Chrysomelidae",        # Crysomelidae -> misspelling
    "Cucujidae",            # Cucujidae
    "Curculionidae",        # Cucurlionidae -> misspelling
    "Curculionidae",        # Culculionidae -> misspelling
    "Curculionidae",        # Curculionidae
    NA,                     # Curculionoidea -> superfamily, not family
    "Dascillidae",          # Dascillidae
    "Melyridae",            # Dasytaeidae -> now part of Melyridae
    "Melyridae",            # Dasytidae -> now part of Melyridae
    "Dermestidae",          # Dermestidae
    "Curculionidae",        # Dryophthoridae -> subfamily of Curculionidae
    "Scarabaeidae",         # Dynastidae -> now subfamily of Scarabaeidae
    "Scarabaeidae",         # Dynastinae -> subfamily of Scarabaeidae
    "Dytiscidae",           # Dytiscidae
    "Elateridae",           # Elateridae
    "Elateridae",           # Elatyridae -> misspelling
    "Erotylidae",           # Erotylidae
    "Eucnemidae",           # Eucmenidae -> misspelling
    "Eucnemidae",           # Eucnemanidae -> misspelling
    "Chrysomelidae",        # Galerucidae -> now subfamily of Chrysomelidae
    "Chrysomelidae",        # Galerucinae -> subfamily of Chrysomelidae
    "Geotrupidae",          # Geotrupidae
    "Glaphyridae",          # Glaphyridae
    "Scirtidae",            # Halodidae -> old name for Scirtidae
    "Scirtidae",            # Helodidae -> old name for Scirtidae
    "Histeridae",           # Histeridae
    "Scarabaeidae",         # Hopliidae -> now part of Scarabaeidae
    "Hybosoridae",          # Hybosoridae
    "Hydrophilidae",        # Hydrophilidae
    "Kateretidae",          # Kateretidae
    "Tenebrionidae",        # Lagridae -> now subfamily of Tenebrionidae
    "Tenebrionidae",        # Lagriidae -> now subfamily of Tenebrionidae
    "Lampyridae",           # Lampyridae
    "Erotylidae",           # Languriidae -> now subfamily of Erotylidae
    "Latridiidae",          # Lathridiidae -> variant spelling
    "Latridiidae",          # Latridiidae
    "Leiodidae",            # Leiodidae
    "Lycidae",              # Lycidae
    "Bostrichidae",         # Lyctidae -> now subfamily of Bostrichidae
    "Lymexylidae",          # Lymexylidae
    "Melyridae",            # Malachidae -> now part of Melyridae
    "Melyridae",            # Malachiidae -> now part of Melyridae
    "Melyridae",            # Malachiinae -> subfamily of Melyridae
    NA,                     # Maloidea -> not a valid family name
    "Megalopodidae",        # Megalopodidae
    "Melyridae",            # Melachiidae -> misspelling
    "Melandryidae",         # Melandryidae
    "Meloidae",             # Meliodae -> misspelling
    "Melyridae",            # Meliridae -> variant spelling
    "Meloidae",             # Meloidae
    "Meloidae",             # Meloididae -> misspelling
    "Scarabaeidae",         # Melolinthinae -> misspelling of Melolonthinae
    "Scarabaeidae",         # Melolonthidae -> now subfamily of Scarabaeidae
    "Scarabaeidae",         # Melolontidae -> misspelling
    "Scarabaeidae",         # Melotonthidae -> misspelling
    "Melyridae",            # Melyridae
    "Monotomidae",          # Monotomidae
    "Mordellidae",          # Mordelidae -> misspelling
    "Mordellidae",          # Mordellidae
    "Mycetophagidae",       # Mycetophagidae
    "Mycteridae",           # Mycteridae
    "Chrysomelidae",        # Mylabridae -> old name for Bruchinae
    "Nitidulidae",          # Nitidulidae
    NA,                     # none
    NA,                     # not specified
    "Oedemeridae",          # Oedemeridae
    "Oedemeridae",          # Oedemeriidae -> misspelling
    "Oedemeridae",          # Oedomeridae -> misspelling
    "Oedemeridae",          # Oemideridae -> misspelling
    "Orsodacnidae",         # Orsodacnidae
    NA,                     # Pentatomidae -> not a beetle family
    "Phalacridae",          # Phalacradidae -> misspelling
    "Phalacridae",          # Phalacridae
    "Phalacridae",          # Phalacrididae -> misspelling
    "Ptiliidae",            # Ptiliidae
    "Ptiliidae",            # Ptilliidae -> misspelling
    "Ptinidae",             # Ptinidae
    "Pythidae",             # Pythidae
    "Ripiphoridae",         # Rhipiphoridae -> variant spelling
    "Monotomidae",          # Rhizophagidae -> now subfamily of Monotomidae
    "Ripiphoridae",         # Ripiphoridae
    "Ripiphoridae",         # Ripriphoridae -> misspelling
    "Scarabaeidae",         # Rutelidae -> now subfamily of Scarabaeidae
    "Scarabaeidae",         # Rutelinae -> subfamily of Scarabaeidae
    "Scarabaeidae",         # Scarabaeidae
    "Scarabaeidae",         # Scarabaeidae (Rutelinae)
    "Scarabaeidae",         # Scarabeidae -> misspelling
    "Scarabaeidae",         # Scarabidae -> misspelling
    NA,                     # Scaraboidea -> superfamily, not family
    "Scirtidae",            # Scirtidae
    "Curculionidae",        # Scolytidae -> now subfamily of Curculionidae
    "Scraptiidae",          # Scraptiidae
    "Staphylinidae",        # Scydmaenidae -> now subfamily of Staphylinidae
    "Oedemeridae",          # Sessinia -> genus in Oedemeridae
    "Staphylinidae",            # Silphidae
    "Silvanidae",           # Silvanidae
    "Silvanidae",           # Silvannidae -> misspelling
    "Staphylinidae",        # Staphilinidae -> misspelling
    "Staphylinidae",        # Staphylinidae
    "Staphylinidae",            # Sylphidae -> misspelling
    "Silvanidae",           # Sylvaniidae -> misspelling
    "Tenebrionidae",        # Tenebrionidae
    "Trogossitidae",        # Trogossitidae
    NA,                     # Undetermined
    NA,                     # Unidentified Coleoptera
    NA,                     # unspecified
    "Zopheridae"            # Zopheridae
  )
)

# Create a function to standardize the mapping
get_accepted_family <- function(original_name) {
  # Convert input to proper case to match our mapping table
  original_name <- str_to_title(original_name)
  
  # Map each name individually
  sapply(original_name, function(name) {
    mapping <- beetle_names %>%
      filter(original_name == name) %>%
      pull(accepted_family)
    
    # Return NA if no mapping found, otherwise return the accepted family name
    if (length(mapping) == 0) NA else mapping[1]
  })
}

# Example usage:
# get_accepted_family("Scolytidae")  # Returns "Curculionidae"
# get_accepted_family("InvalidName") # Returns NA
beetle_names
```

Now let's apply the beetle family name mapping:
```{r}
plant_family_output_df = references_for_plant_families %>%
    rowwise() %>%
    mutate(
        beetle_families = list({
            if(is.null(beetle_families)) {
                NULL
            } else if(length(unlist(beetle_families)) > 0) {
                families <- unlist(beetle_families)
                accepted <- get_accepted_family(families)
                unique(accepted)
            } else {
                character(0)
            }
        })
    )

plant_family_output_df
```
How many plant families total have had beetle flower visitors recorded?
```{r}
plant_family_output_df %>%
  filter(beetle_visitors | beetle_pollinators) %>%
  pull(plant_family) %>%
  unique %>%
  length()
```

How many plant families total do we have?
```{r}
plant_family_output_df %>%
  pull(plant_family) %>%
  unique %>%
  length()
```

How many plant families total have at least one paper mentioning beetle pollinators?
```{r}
plant_family_output_df %>%
  filter(beetle_pollinators) %>%
  pull(plant_family) %>%
  unique %>%
  length()
```

What are those families?
```{r}
plant_family_output_df %>%
  filter(beetle_pollinators) %>%
  rowwise() %>%
  mutate(N_references = length(unlist(bibtex_ids))) %>%
  ungroup() %>%
  arrange(desc(N_references))
```

What are the top families in beetle diversity?
```{r}
plant_family_output_df %>%
  filter(!is.null(beetle_families)) %>%
  group_by(plant_family) %>%
  summarise(N_beetles=length(unlist(beetle_families))) %>%
  ungroup() %>%
  arrange(desc(N_beetles))
```
What are the top families in number of records with beetles?
```{r}
filtered_state_level_records %>% filter(beetle_visitors) %>%
  group_by(plant_family_name) %>%
  summarise(N_records = n()) %>%
  ungroup() %>%
  arrange(desc(N_records))
```


# Analising beetle family occurence and co-occurence patterns
We will now analyse co-occurence patterns in beetle families and whether they are related to pollination. Let's first create the dataframe we will need. In this case, we will only keep the records in which at least one beetle family has been identified:
```{r}
df_for_pcoa <- filtered_state_level_records %>%
  # Create and recode beetle status
  mutate(
    beetle_status = case_when(
      beetle_visitors & beetle_pollinators ~ 'pollinators',
      beetle_visitors ~ 'nonpollinators',
      TRUE ~ 'none'
    ),
    beetle_status = factor(beetle_status, levels = c("pollinators", "nonpollinators", "none")),
    record_id = 1:n()
  ) %>%
  
  # Process beetle families
  rowwise() %>%
      mutate(
        beetle_families = list({
            if(is.null(beetle_families)) {
                NULL
            } else if(length(unlist(beetle_families)) > 0) {
                families <- unlist(beetle_families)
                accepted <- get_accepted_family(families)
                unique(accepted)
            } else {
                character(0)
            }
        })
    ) %>%
  
  # Create presence/absence matrix
  select(record_id, plant_family_name, beetle_status, beetle_families) %>%
  unnest(beetle_families) %>%
  mutate(present = TRUE) %>%
  pivot_wider(
    names_from = beetle_families,
    values_from = present,
    values_fill = FALSE
  )

df_for_pcoa
```

Let's use this now to count the number of records for each family:
```{r}
df_for_pcoa %>% 
  select(-record_id, -plant_family_name, -beetle_status) %>% 
  ungroup() %>%
  summarise_all(sum) %>%
  pivot_longer(cols = everything(), names_to = "family",values_to = "N_records") %>%
  arrange(desc(N_records)) %>%
  mutate(p_records = N_records/nrow(filtered_state_level_records))
```

Let's now do the PCoA and plot:
```{r}
# Create distance matrix using only beetle family columns
beetle_matrix <- df_for_pcoa %>%
  select(-record_id, -plant_family_name, -beetle_status) %>%
  as.matrix()

# Perform PCoA using Jaccard distance
pcoa <- cmdscale(vegdist(beetle_matrix, method = "jaccard"), k = 2, eig = TRUE)

# Create points dataframe
points_df <- data.frame(
  PCoA1 = pcoa$points[,1],
  PCoA2 = pcoa$points[,2],
  beetle_status = df_for_pcoa$beetle_status
)

# Calculate species scores for ALL families
species_scores <- cor(beetle_matrix, pcoa$points)
all_arrows_df <- data.frame(
  Family = colnames(beetle_matrix),
  PCoA1 = species_scores[,1],
  PCoA2 = species_scores[,2]
) %>%
  mutate(vector_length = sqrt(PCoA1^2 + PCoA2^2))

# Get top 10 for labeled arrows
arrows_df <- all_arrows_df %>%
  arrange(desc(vector_length)) %>%
  slice_head(n = 8)

# Calculate proportion of variance explained
var_explained <- round(100 * pcoa$eig / sum(pcoa$eig), 2)

# Create the plot
pcoa_plot = ggplot() +
  # Add light arrows for all families
  geom_jitter(data = points_df, 
             aes(x = PCoA1, y = PCoA2, color = beetle_status),
             size = 0.5,
             alpha = 0.6,
             width = 0.01,
             height = 0.01) +
  # Add darker arrows for top 10 families
  geom_segment(data = all_arrows_df,
               aes(x = 0, y = 0, xend = PCoA1, yend = PCoA2),
               arrow = arrow(length = unit(0.1, "cm")),
               size = 0.3,
               color = "gray60") +
  geom_segment(data = arrows_df,
               aes(x = 0, y = 0, xend = PCoA1, yend = PCoA2),
               arrow = arrow(length = unit(0.1, "cm")),
               size = 0.3,
               color = "gray30") +
  geom_text_repel(data = arrows_df,
                  aes(x = PCoA1, y = PCoA2, label = Family),
                  size = 1.8,
                  force_pull = 10,
                  force = 1,
                  max.overlaps = Inf,
                  box.padding = 0.001,
                  point.padding = 0.01,
                  segment.color = NA) +
  scale_color_brewer(palette = "Set2") +
  labs(x = paste0("PCoA1 (", var_explained[1], "%)"),
       y = paste0("PCoA2 (", var_explained[2], "%)"),
       color = "Beetle Status") +
  coord_equal() +
  theme_minimal() +
  guides(fill = guide_legend(title.position = "top",
    title.hjust = 0.5)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        text = element_text(size=8),
               legend.position = "bottom",
    legend.box = "horizontal",
    legend.key.size = unit(0.4, "lines"),
    legend.spacing.x = unit(0.1, "cm"),
    legend.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt")
        )

pcoa_plot
```

Let's now combine this plot with the others:
```{r}
combined_nongeo <- patchwork::wrap_plots(
  p_bars,
  p_years + guides(fill='none'),
  pcoa_plot + guides(fill='none') + coord_cartesian(),
  ncol = 1,
  heights = c(1,0.8,1)
) +
  plot_layout(guides = 'collect') +  # This collects and combines similar legends
  plot_annotation(
    theme = theme(
      legend.position = 'bottom',  # Put all legends on the right
    )
  )

combined_nongeo
ggsave(combined_nongeo, filename = 'figures/fig1BCD.pdf',device = 'pdf',units = 'mm', height = 180, width=70, useDingbats=F)
```




# Plotting beetle phylogeny

We will now create a visualization of beetle phylogeny highlighting flower-visiting families. The visualization is based on:

1. A phylogenetic tree from `Cai_et_al_phylo/FigTree.tre` (from Cai et al, 2022)
2. Taxonomic classification from the supplement in Cai et al, 2022
3. Data on flower-visiting behavior from our review (Supplementary Table 1)


## Create Taxonomic Database

We will create a data frame of beetle taxonomy, including suborders, superfamilies, and families from Cai et al (2022). We'll now add information about which families are known to visit flower from our Supplementary Table 1.

```{r taxonomy}
beetle_taxonomy <- tribble(
  ~Suborder, ~Superfamily, ~Family,
  # ADEPHAGA
  "Adephaga", "Caraboidea", "Carabidae",  
  "Adephaga", "Caraboidea", "Cicindelidae",
  "Adephaga", "Caraboidea", "Trachypachidae",
  "Adephaga", "Dytiscoidea", "Amphizoidae",
  "Adephaga", "Dytiscoidea", "Aspidytidae",
  "Adephaga", "Dytiscoidea", "Dytiscidae",
  "Adephaga", "Dytiscoidea", "Hygrobiidae",
  "Adephaga", "Dytiscoidea", "Meruidae",
  "Adephaga", "Dytiscoidea", "Noteridae",
  "Adephaga", "Gyrinoidea", "Gyrinidae",
  "Adephaga", "Haliploidea", "Haliplidae",
  
  # ARCHOSTEMATA
  "Archostemata", "Cupedoidea", "Crowsoniellidae",
  "Archostemata", "Cupedoidea", "Cupedidae",
  "Archostemata", "Cupedoidea", "Micromalthidae",
  "Archostemata", "Cupedoidea", "Ommatidae",
  
  # MYXOPHAGA
  "Myxophaga", "Lepiceroidea", "Lepiceridae",
  "Myxophaga", "Sphaeriusoidea", "Hydroscaphidae",
  "Myxophaga", "Sphaeriusoidea", "Sphaeriusidae",
  "Myxophaga", "Sphaeriusoidea", "Torridincolidae",
  
  # POLYPHAGA
  "Polyphaga", "Bostrichoidea", "Bostrichidae",
  "Polyphaga", "Bostrichoidea", "Dermestidae",
  "Polyphaga", "Bostrichoidea", "Ptinidae",
  
  "Polyphaga", "Buprestoidea", "Buprestidae",
  "Polyphaga", "Buprestoidea", "Schizopodidae",
  
  "Polyphaga", "Byrrhoidea", "Byrrhidae",
  
  "Polyphaga", "Dryopoidea", "Callirhipidae",
  "Polyphaga", "Dryopoidea", "Chelonariidae",
  "Polyphaga", "Dryopoidea", "Dryopidae",
  "Polyphaga", "Dryopoidea", "Elmidae",
  "Polyphaga", "Dryopoidea", "Eulichadidae",
  "Polyphaga", "Dryopoidea", "Heteroceridae",
  "Polyphaga", "Dryopoidea", "Limnichidae",
  "Polyphaga", "Dryopoidea", "Lutrochidae",
  "Polyphaga", "Dryopoidea", "Protelmidae",
  "Polyphaga", "Dryopoidea", "Psephenidae",
  "Polyphaga", "Dryopoidea", "Ptilodactylidae",
  "Polyphaga", "Dryopoidea", "Cneoglossidae",


  
  "Polyphaga", "Chrysomeloidea", "Cerambycidae",
  "Polyphaga", "Chrysomeloidea", "Chrysomelidae",
  "Polyphaga", "Chrysomeloidea", "Disteniidae",
  "Polyphaga", "Chrysomeloidea", "Megalopodidae",
  "Polyphaga", "Chrysomeloidea", "Orsodacnidae",
  "Polyphaga", "Chrysomeloidea", "Oxypeltidae",
  "Polyphaga", "Chrysomeloidea", "Vesperidae",
  
  "Polyphaga", "Clamboidea", "Clambidae",
  "Polyphaga", "Clamboidea", "Derodontidae",
  "Polyphaga", "Clamboidea", "Eucinetidae",
  
  "Polyphaga", "Cleroidea", "Acanthocnemidae",
  "Polyphaga", "Cleroidea", "Biphyllidae",
  "Polyphaga", "Cleroidea", "Byturidae",
  "Polyphaga", "Cleroidea", "Chaetosomatidae",
  "Polyphaga", "Cleroidea", "Cleridae",
  "Polyphaga", "Cleroidea", "Lophocateridae",
  "Polyphaga", "Cleroidea", "Mauroniscidae",
  "Polyphaga", "Cleroidea", "Melyridae",
  "Polyphaga", "Cleroidea", "Peltidae",
  "Polyphaga", "Cleroidea", "Phloiophilidae",
  "Polyphaga", "Cleroidea", "Phycosecidae",
  "Polyphaga", "Cleroidea", "Prionoceridae",
  "Polyphaga", "Cleroidea", "Protopeltidae",
  "Polyphaga", "Cleroidea", "Rentoniidae",
  "Polyphaga", "Cleroidea", "Rhadalidae",
  "Polyphaga", "Cleroidea", "Thanerocleridae",
  "Polyphaga", "Cleroidea", "Thymalidae",
  "Polyphaga", "Cleroidea", "Trogossitidae",
  
  "Polyphaga", "Coccinelloidea", "Akalyptoischiidae",
  "Polyphaga", "Coccinelloidea", "Alexiidae",
  "Polyphaga", "Coccinelloidea", "Anamorphidae",
  "Polyphaga", "Coccinelloidea", "Bothrideridae",
  "Polyphaga", "Coccinelloidea", "Cerylonidae",
  "Polyphaga", "Coccinelloidea", "Coccinellidae",
  "Polyphaga", "Coccinelloidea", "Corylophidae",
  "Polyphaga", "Coccinelloidea", "Discolomatidae",
  "Polyphaga", "Coccinelloidea", "Endomychidae",
  "Polyphaga", "Coccinelloidea", "Eupsilobiidae",
  "Polyphaga", "Coccinelloidea", "Euxestidae",
  "Polyphaga", "Coccinelloidea", "Latridiidae",
  "Polyphaga", "Coccinelloidea", "Murmidiidae",
  "Polyphaga", "Coccinelloidea", "Mycetaeidae",
  "Polyphaga", "Coccinelloidea", "Teredidae",
  
  "Polyphaga", "Cucujoidea", "Agapythidae",
  "Polyphaga", "Cucujoidea", "Cavognathidae",
  "Polyphaga", "Cucujoidea", "Cryptophagidae",
  "Polyphaga", "Cucujoidea", "Cucujidae",
  "Polyphaga", "Cucujoidea", "Cyclaxyridae",
  "Polyphaga", "Cucujoidea", "Hobartiidae",
  "Polyphaga", "Cucujoidea", "Laemophloeidae",
  "Polyphaga", "Cucujoidea", "Lamingtoniidae",
  "Polyphaga", "Cucujoidea", "Myraboliidae",
  "Polyphaga", "Cucujoidea", "Passandridae",
  "Polyphaga", "Cucujoidea", "Phalacridae",
  "Polyphaga", "Cucujoidea", "Phloeostichidae",
  "Polyphaga", "Cucujoidea", "Priasilphidae",
  "Polyphaga", "Cucujoidea", "Prostomidae",
  "Polyphaga", "Cucujoidea", "Silvanidae",
  "Polyphaga", "Cucujoidea", "Tasmosalpingidae",
  
  "Polyphaga", "Curculionoidea", "Anthribidae",
  "Polyphaga", "Curculionoidea", "Attelabidae",
  "Polyphaga", "Curculionoidea", "Belidae",
  "Polyphaga", "Curculionoidea", "Brentidae",
  "Polyphaga", "Curculionoidea", "Caridae",
  "Polyphaga", "Curculionoidea", "Cimberididae",
  "Polyphaga", "Curculionoidea", "Curculionidae",
  "Polyphaga", "Curculionoidea", "Nemonychidae",
  
  "Polyphaga", "Dascilloidea", "Dascillidae",
  "Polyphaga", "Dascilloidea", "Rhipiceridae",
  
  "Polyphaga", "Elateroidea", "Artematopodidae",
  "Polyphaga", "Elateroidea", "Brachypsectridae",
  "Polyphaga", "Elateroidea", "Cantharidae",
  "Polyphaga", "Elateroidea", "Cerophytidae",
  "Polyphaga", "Elateroidea", "Elateridae",
  "Polyphaga", "Elateroidea", "Eucnemidae",
  "Polyphaga", "Elateroidea", "Iberobaeniidae",
  "Polyphaga", "Elateroidea", "Jurasaidae",
  "Polyphaga", "Elateroidea", "Lampyridae",
  "Polyphaga", "Elateroidea", "Lycidae",
  "Polyphaga", "Elateroidea", "Omethidae",
  "Polyphaga", "Elateroidea", "Phengodidae",
  "Polyphaga", "Elateroidea", "Rhagophthalmidae",
  "Polyphaga", "Elateroidea", "Sinopyrophoridae",
  "Polyphaga", "Elateroidea", "Throscidae",
  
  "Polyphaga", "Erotyloidea", "Boganiidae",
  "Polyphaga", "Erotyloidea", "Erotylidae",
  
  "Polyphaga", "Histeroidea", "Histeridae",
  "Polyphaga", "Histeroidea", "Sphaeritidae",
  "Polyphaga", "Histeroidea", "Synteliidae",
  
  "Polyphaga", "Hydrophiloidea", "Epimetopidae",
  "Polyphaga", "Hydrophiloidea", "Georissidae",
  "Polyphaga", "Hydrophiloidea", "Helophoridae",
  "Polyphaga", "Hydrophiloidea", "Hydrochidae",
  "Polyphaga", "Hydrophiloidea", "Hydrophilidae",
  "Polyphaga", "Hydrophiloidea", "Spercheidae",
  
  "Polyphaga", "Lymexyloidea", "Lymexylidae",
  
  "Polyphaga", "Nitiduloidea", "Helotidae",
  "Polyphaga", "Nitiduloidea", "Kateretidae",
  "Polyphaga", "Nitiduloidea", "Monotomidae",
  "Polyphaga", "Nitiduloidea", "Nitidulidae",
  "Polyphaga", "Nitiduloidea", "Protocucujidae",
  "Polyphaga", "Nitiduloidea", "Sphindidae",
  "Polyphaga", "Nitiduloidea", "Smicripidae",


  
  "Polyphaga", "Nosodendroidea", "Nosodendridae",
  
  "Polyphaga", "Rhinorhipoidea", "Rhinorhipidae",
  
  "Polyphaga", "Scarabaeoidea", "Bolboceratidae",
  "Polyphaga", "Scarabaeoidea", "Belohinidae",
  "Polyphaga", "Scarabaeoidea", "Diphyllostomatidae",
  "Polyphaga", "Scarabaeoidea", "Glaphyridae",
  "Polyphaga", "Scarabaeoidea", "Glaresidae",
  "Polyphaga", "Scarabaeoidea", "Geotrupidae",
  "Polyphaga", "Scarabaeoidea", "Hybosoridae",
  "Polyphaga", "Scarabaeoidea", "Lucanidae",
  "Polyphaga", "Scarabaeoidea", "Ochodaeidae",
  "Polyphaga", "Scarabaeoidea", "Passalidae",
  "Polyphaga", "Scarabaeoidea", "Pleocomidae",
  "Polyphaga", "Scarabaeoidea", "Scarabaeidae",
  "Polyphaga", "Scarabaeoidea", "Trogidae",
  
  "Polyphaga", "Scirtoidea", "Decliniidae",
  "Polyphaga", "Scirtoidea", "Scirtidae",
  
  "Polyphaga", "Staphylinoidea", "Agyrtidae",
  "Polyphaga", "Staphylinoidea", "Colonidae",
  "Polyphaga", "Staphylinoidea", "Hydraenidae",
  "Polyphaga", "Staphylinoidea", "Jacobsoniidae",
  "Polyphaga", "Staphylinoidea", "Leiodidae",
  "Polyphaga", "Staphylinoidea", "Ptiliidae",
  "Polyphaga", "Staphylinoidea", "Staphylinidae",
  
  "Polyphaga", "Tenebrionoidea", "Aderidae",
  "Polyphaga", "Tenebrionoidea", "Anthicidae",
  "Polyphaga", "Tenebrionoidea", "Archeocrypticidae",
  "Polyphaga", "Tenebrionoidea", "Boridae",
  "Polyphaga", "Tenebrionoidea", "Chalcodryidae",
  "Polyphaga", "Tenebrionoidea", "Ciidae",
  "Polyphaga", "Tenebrionoidea", "Ischaliidae",
  "Polyphaga", "Tenebrionoidea", "Melandryidae",
  "Polyphaga", "Tenebrionoidea", "Meloidae",
  "Polyphaga", "Tenebrionoidea", "Mordellidae",
  "Polyphaga", "Tenebrionoidea", "Mycetophagidae",
  "Polyphaga", "Tenebrionoidea", "Mycteridae",
  "Polyphaga", "Tenebrionoidea", "Oedemeridae",
  "Polyphaga", "Tenebrionoidea", "Promecheilidae",
  "Polyphaga", "Tenebrionoidea", "Pterogeniidae",
  "Polyphaga", "Tenebrionoidea", "Pyrochroidae",
  "Polyphaga", "Tenebrionoidea", "Pythidae",
  "Polyphaga", "Tenebrionoidea", "Ripiphoridae",
  "Polyphaga", "Tenebrionoidea", "Salpingidae",
  "Polyphaga", "Tenebrionoidea", "Scraptiidae",
  "Polyphaga", "Tenebrionoidea", "Stenotrachelidae",
  "Polyphaga", "Tenebrionoidea", "Synchroidae",
  "Polyphaga", "Tenebrionoidea", "Tenebrionidae",
  "Polyphaga", "Tenebrionoidea", "Tetratomidae",
  "Polyphaga", "Tenebrionoidea", "Trictenotomidae",
  "Polyphaga", "Tenebrionoidea", "Ulodidae",
  "Polyphaga", "Tenebrionoidea", "Zopheridae",
  
  # Polyphaga incertae sedis
  "Polyphaga", "incertae sedis", "Jurodidae"
)

# Add flower visitor information
flower_visitors <- c("Agyrtidae", "Anthicidae", "Anthribidae", "Attelabidae", "Belidae", "Biphyllidae", "Boganiidae", "Bostrichidae", "Byrrhidae", "Brentidae", "Byturidae", "Buprestidae", "Cantharidae", "Carabidae", "Cerambycidae", "Chrysomelidae", "Cicindelidae", "Cleridae", "Coccinellidae", "Corylophidae", "Cryptophagidae", "Cucujidae", "Curculionidae", "Dascillidae", "Dermestidae", "Dytiscidae", "Elateridae", "Erotylidae", "Eucnemidae", "Geotrupidae", "Glaphyridae", "Histeridae", "Hydrophilidae", "Hybosoridae", "Kateretidae", "Laemophloeidae", "Lampyridae", "Latridiidae", "Leiodidae", "Lymexylidae", "Lucanidae", "Lycidae", "Megalopodidae", "Meloidae", "Melyridae", "Monotomidae", "Mordellidae", "Mycetophagidae", "Mycteridae", "Nemonychidae", "Nitidulidae", "Oedemeridae", "Orsodacnidae", "Phalacridae", "Ptinidae", "Ptiliidae", "Pyrochroidae", "Ripiphoridae", "Salpingidae", "Scarabaeidae", "Scraptiidae", "Scirtidae", "Silvanidae", "Sphaeroceridae", "Staphylinidae", "Stenotrachelidae", "Tenebrionidae", "Throscidae", "Trogossitidae", "Zopheridae")

pollinator_families <- c("Anthicidae", "Anthribidae", "Attelabidae", "Belidae", "Biphyllidae", "Boganiidae", "Brentidae", "Byturidae", "Buprestidae", "Cantharidae", "Carabidae", "Cerambycidae", "Chrysomelidae", "Cleridae", "Coccinellidae", "Corylophidae", "Cryptophagidae", "Curculionidae", "Dermestidae", "Elateridae", "Erotylidae", "Geotrupidae", "Glaphyridae", "Hybosoridae", "Hydrophilidae", "Kateretidae", "Lampyridae", "Latridiidae", "Lymexylidae", "Lucanidae", "Lycidae", "Meloidae", "Melyridae", "Monotomidae", "Mordellidae", "Mycetophagidae", "Nitidulidae", "Oedemeridae", "Ptinidae", "Ptiliidae", "Pyrochroidae", "Scarabaeidae", "Scraptiidae", "Scirtidae", "Silvanidae", "Sphaeroceridae", "Staphylinidae", "Stenotrachelidae", "Tenebrionidae", "Throscidae")

beetle_taxonomy$flower_visitor <- beetle_taxonomy$Family %in% flower_visitors
beetle_taxonomy$pollinator <- beetle_taxonomy$Family %in% pollinator_families
```

## Process Phylogenetic Tree

Next, we'll load and process the phylogenetic tree, removing outgroups and simplifying monophyletic groups.

```{r tree_processing}
# Read tree and remove outgroups
tree <- read.nexus("Cai_et_al_phylo/FigTree.tre")
outgroups <- c("Corydalidae", "Chrysopidae", "Myrmeleontidae", "Ascalaphidae")
tips_to_keep <- tree$tip.label[!grepl(paste(outgroups, collapse="|"), tree$tip.label)]
tree <- keep.tip(tree, tips_to_keep)
```

Update Silphidae and Cybocephalidae status
```{r}
tree$tip.label = str_replace(tree$tip.label, "Silphidae", "Staphylinidae") %>%
  str_replace("Cybocephalidae", "Nitidulidae")
```


Collapse monophyletic family-name groups
```{r}
while(TRUE) {
  changes_made <- FALSE
  N <- Ntip(tree)
  internal_nodes <- (N + 1):(N + tree$Nnode)
  
  for(node in internal_nodes) {
    descendants <- extract.clade(tree, node)$tip.label
    families <- unique(gsub("_.*$", "", descendants))
    
    if(length(families) == 1 && length(descendants) > 1) {
      # Keep first tip, drop others
      tips_to_drop <- descendants[-1]
      tree <- drop.tip(tree, tips_to_drop)
      changes_made <- TRUE
      break  # Break the for loop as node numbers have changed
    }
  }
  
  if(!changes_made) break  # No more changes needed, exit while loop
}
```

Rename tips
```{r}
# First ladderize the tree
tree <- ladderize(tree)

# Get family names
families <- gsub("_.*$", "", tree$tip.label)

# Create new labels
new_labels <- character(length(tree$tip.label))
family_counts <- table(families)

# For each unique family
for(family in unique(families)) {
  # If family appears only once, use simple name
  if(family_counts[family] == 1) {
    new_labels[families == family] <- family
  } else {
    # If family appears multiple times, add numbering
    new_labels[families == family] <- paste(family, 1:family_counts[family])
  }
}

# Replace tip labels
tree$tip.label <- new_labels
```


## Find Taxonomic Nodes

First,we will update our taxonomic dataframe to relate families to tip labels. 
```{r}
# Get the tip labels and their corresponding families
tip_data <- tibble(
  node = 1:Ntip(tree),
  tip_label = tree$tip.label,
  Family = gsub(" \\d+$", "", tree$tip.label)  # Remove numbering from family names
) %>%
  # Add flower visitor information
  mutate(is_flower_visitor = Family %in% unique(
    beetle_taxonomy$Family[beetle_taxonomy$flower_visitor]
  ),
  is_pollinator = Family %in% unique(
    beetle_taxonomy$Family[beetle_taxonomy$pollinator]
  ),
  beetle_status = case_when(
   is_pollinator == TRUE ~ "pollinator",
   is_flower_visitor == TRUE ~ "flower visitor",
   TRUE ~ NA
 ))

# Function to find tip labels for a family
get_tip_labels <- function(family, tip_data) {
  matches <- tip_data$tip_label[tip_data$Family == family]
  if(length(matches) == 0) return(NA)
  return(matches)
}

# Create new data frame with duplicated rows for multiple tips
new_beetle_taxonomy <- do.call(rbind, lapply(1:nrow(beetle_taxonomy), function(i) {
  family <- beetle_taxonomy$Family[i]
  tip_labels <- get_tip_labels(family, tip_data)
  
  if(length(tip_labels) == 1 && is.na(tip_labels)) {
    # Family not in tree
    return(cbind(beetle_taxonomy[i,], tip_label = NA))
  } else {
    # Create one row per tip label
    return(do.call(rbind, lapply(tip_labels, function(tip) {
      cbind(beetle_taxonomy[i,], tip_label = tip)
    })))
  }
}))

# Convert back to tibble if needed
beetle_taxonomy <- as_tibble(new_beetle_taxonomy)
```

These are the flower visitor families missing from the tree:
```{r}
beetle_taxonomy %>% filter(flower_visitor, is.na(tip_label))
```


Now we'll identify the nodes corresponding to suborders and superfamilies.

## Create Visualization

Finally, we'll create the visualization with colored branches and taxonomic labels.
```{r}
# First rescale the tree to geological time
tree$edge.length <- tree$edge.length * 100  # Multiply by 100 to get ~300 MY depth
```


```{r plot_tree, fig.width=12, fig.height=16}
# Create base plot
p <- ggtree(tree) + 
  # Add gray branches
  geom_tree(size = 0.1, color = "gray20") +
  # Add points for flower visitors
  geom_point(
    data = . %>% filter(isTip & tip_data$is_flower_visitor[node] &!tip_data$is_pollinator[node]), 
    aes(x = x, y = y),
    color = "goldenrod",
    size = 1
  ) +
  geom_point(
    data = . %>% filter(isTip & tip_data$is_pollinator[node]), 
    aes(x = x, y = y),
    color = "brown1",
    size = 1
  ) 

# First get the superfamily ranges
superfamily_ranges <- p$data %>% 
  filter(isTip) %>% 
  left_join(beetle_taxonomy, by = c('label'='tip_label')) %>% 
  group_by(Superfamily) %>% 
  summarise(min_y = min(y), max_y = max(y)) %>% 
  mutate(min_y = min_y - 0.4, max_y = max_y + 0.4) %>%
  filter(!is.na(Superfamily))

# Create color palette for superfamilies
superfamily_colors <- setNames(
  scales::hue_pal()(length(unique(beetle_taxonomy$Superfamily))),
  unique(beetle_taxonomy$Superfamily)
)

# Create base plot
p <- p + 
  # Add rectangles for each superfamily
  geom_rect(
    data = superfamily_ranges,
    aes(
      ymin = min_y,
      ymax = max_y,
      xmin = 10,
      xmax = 80,
      fill = Superfamily
    ),
    inherit.aes = FALSE,
    alpha = 0.7
  ) +
  # Add superfamily labels inside rectangles
  geom_text(
    data = superfamily_ranges,
    aes(
      x = 45,  # Center of rectangle
      y = (min_y + max_y)/2,  # Center of rectangle
      label = Superfamily
    ),
    color = 'black',
    size = 1.5,
    hjust = 0.5,
  ) +
  scale_fill_manual(values = superfamily_colors, guide = 'none') +
  scale_color_manual(values = superfamily_colors, guide = 'none') +
  coord_geo(
    dat = "periods",
    pos = "bottom", 
    xlim = c(-330,100),
    ylim = c(0,142),
    neg = TRUE,
    height = unit(0.8, "lines"),
    size = 3,
    center_end_labels = TRUE
  ) +
  scale_x_continuous(
  breaks = seq(-300, 0, by = 50),
  labels = abs(seq(300, 0, by = -50)),
  
) +
  theme_tree2()

revts(p)

ggsave(filename = 'figures/Coleop_tree.pdf', device = 'pdf', useDingbats=F, height = 180, width = 100, units = 'mm' )
```


# Session Info

```{r session_info}
sessionInfo()
```







